\textbf{\LARGE Time Is Running Out\\ \large Assessing Temporal Privacy of Privacy Zones in Fitness Tracking Social Networks}

### 2023-06-24 article.tex
Verandering van volgorde ch2

Fitness trackers each implement ways to improve user privacy. Perhaps the most
easy way to think of is perhaps the ability to hide activities from a selection
of people (e.g., anyone who is not a follower). Thus, only the people whom the
user explicitly allows to view activities. A more complex alternative is to use
EPZs. Hereby will a portion of the displayed route be hidden for external
viewers. Part of the route is cut off, so to speak. The real starting and
ending points will lie inside the truncated part. New points will be generated,
on the edge of the circle, which for the external observer will represent the
beginning and end. The beginning and end part of the route will thus become
invisible to the other users. Due to the presence of all these attempts at
privacy enhancements, it is noticeable that the developers of the platforms are
very aware of the potential dangers. However, there is a trade-off to be made
when implementing between the usability of the platform, and the privacy of the
end user. The more data is released, the greater the chance of potentially
sensitive info being passed along. On the other hand, when omitting
information, the usability and presence of useful data of the platform is
greatly diminished.

@@

To achieve this objective, we must first examine the attack model according to
Dhondt el al. Then we can describe different possible alternative attack
scenarios, and for each work out the calculations necessary in order for this
scenario to be possible. Before we can test and analyze the attack, we must
consider the possible errors in the used dataset. We also perform an analysis
on the difference between the calculated distances, and the values derived
according to the calculations of Dhondt et al. So can we estimate the
effectiveness of the attack a priori to some extent. Only then do we execute
the attack, evaluate the attacks and form meaningful conclusions.

@@

\subsection{GPS faults}
Some visible data on an activity, which can also be found on
Figure~\ref{fig:activity}, are the distance, the duration, the average speed,
etc. The average speed forms the core to the attack which we will describe in
this article. This will be calculated as the total distance divided by the
moving time of the athlete. Fitness trackers receive raw data from the devices.
This data must therefore be processed before it is usefull fo the users.
Especially GPS-data, which is can be exposed to a lot of faults and noise.
There are three types of GPS faults, namely \textit{GPS drift}, \textit{GPS
    signal loss} and \textit{GPS bounce}. GPS-drift is a phenomenon where a user's
GPS location deviates from the effective location. This may be caused by
densely built environments, and natural factors such as tall trees. GPS
bouncing is a phenomenon caused mainly by tall buildings. In this situation,
the GPS signal will bounce in between buildings on its way to the device from
the satellite. The extra delay that the bounces bring along causes the device
to mistakenly think it has traveled some additional distance. The outcome of
the trajectory is then unpredictable, leading to a `cluster' of GPS points. A
last phenomenon that can occur is GPS signal loss. This occurs when the user's
signal is lost, and a new signal is only received at a later time stamp,
causing a jump. A second cause that can lead to signal loss, which especially
applies to fitness trackers, is the ability to pause an activity. I When the
activity is resumed again, there will be a jump in GPS locations, which may
lead to miscalculation of distance.

@@

Some techniques can be used to improve the overall accuracy of the GPS data,
and so improve the effectiveness of the attack. The main technique used in this
thesis is \textit{smoothing}, and the hypothesis is that fitnessplatformen use
this technique as well. This is a technique to enhance the accuracy and
reliability of GPS data. There are different implementations of GPS smoothing,
but the one used in this context is moving average filtering. This method
calculates an average position by considering a sliding window of the most
recent GPS measurements. By averaging multiple measurements over a certain time
period, the effects of noise and temporary inaccuracies can be mitigated,
resulting in a smoother and more reliable trajectory. The size of the window
can be chosen. The larger the window, the less accurate the trajectory will be,
but the more noise will be countered. This technique is especially useful to
counter GPS bouncing and GPS drift.

@@

    \item Then can we move on to the identification of Entry Gates (E.G.). Entry gates

@@

          smaller than the maximum possible distance to be covered. Similarly, filtering

@@

\subsection{Geographical distribution}
A geographical distribution is visible in Figure~\ref{fig:heatmap} It clearly
shows that most activities are located in Central Europe. Additionally, there
is a noticeable concentration in the United States. To a lesser extent, there
are also activities in Australia and South America. The dataset exhibits a
relatively broad spread of activities worldwide, providing a solid foundation
for testing the attack. However, it is important to note that the fraction of
the dataset we have access to, with 101 users, is relatively small, which may
result in a distorted representation of reality.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\linewidth]{fig/Afwijkingen&Analyses/Heatmap.png}\label{fig:heatmap}
    \caption{Geo-heatmap of the users in the dataset}
\end{figure}

\subsection{Users and activities}
Table~\ref{tab:stats_dataset} displays several global statistics regarding
users and their associated activities in the dataset. Figure
~\ref{fig:cdf_amount_activities} shows the CDF plot, illustrating the number of
activities per user. It is notable that there is a significant number of
activities available per user in the dataset. However, it's important to note
that the dataset, with an average of 411 activities per user, is not entirely
representative of reality. When comparing these figures with data from a study
conducted by Strava itself in 2020, there is a mismatch apparent.
\begin{table}[h]
    \centering
    \begin{tabular}{|l||c|}
        \hline
                                                         & \textbf{Aantal} \\
        \hline \hline
        Totaal \# gebruikers                             & 101             \\
        \hline
        Totaal aantal activiteiten                       & 41 554          \\
        \hline
        Gemiddeld \# activiteiten per gebruiker          & 411             \\
        \hline
        Mediaan van het \# activiteiten per gebruiker    & 296             \\
        \hline
        Maximaal \# activities voor een enkele gebruiker & 2946            \\
        \hline
        Minimaal \# activities voor een enkele gebruiker & 31              \\
        \hline
    \end{tabular}
    \captionsetup{justification=centering}
    \caption{Overview of users and activities}\label{tab:stats_dataset}
\end{table}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{fig/Afwijkingen&Analyses/CDF_amountActivities.jpg}
    \caption{CDF plot of the amount of activities per user}\label{fig:cdf_amount_activities}
\end{figure}

@@

The \textit{Certainty metric} quantifies the concentration of the probability
distribution. A higher Certainty value indicates that the nodes in the
probability distribution are spread farther apart, indicating a greater level
of certainty or confidence in the predictions. Essentially, it measures the
degree of dispersion or concentration of the predicted locations in the
probability distribution.

\textit{Spatial Certainty} is a metric that considers the density or concentration of
neighboring nodes in the vicinity of each predicted node, instead of focusing
solely on the probability of each individual node. It takes into account the
density or clustering of predicted locations in the neighborhood of each node.
A higher Spatial Certainty value indicates a higher density or concentration of
predicted nodes in the local vicinity, suggesting a higher level of certainty
in the predicted locations within their respective neighborhoods.

>>>>

Fitness trackers each implement ways to improve user privacy. Perhaps the most
easy way to think of is perhaps the ability to hide activities from a selection
of people (e.g., anyone who is not a follower). Thus, only the people whom the
user explicitly allows to view activities. A more complex alternative is to use
EPZs. Hereby a portion of the displayed route will be hidden for external
viewers. Part of the route is cut off, so to speak. The real starting and
ending points will lie inside the truncated part. New points will be generated,
on the edge of the circle, which for the external observer will represent the
beginning and end. The beginning and end part of the route will thus become
invisible to the other users. Due to the presence of all these attempts at
privacy enhancements, it is noticeable that the developers of the platforms are
very aware of the potential dangers. However, there is a trade-off to be made
when implementing between the usability of the platform, and the privacy of the
end user. The more data is released, the greater the chance of potentially
sensitive info being passed along. On the other hand, when omitting
information, the usability and presence of useful data of the platform is
greatly diminished.

@@

To achieve this objective, we must first examine the attack model according to
Dhondt et al. Then we can describe different possible alternative attack
scenarios, and for each work out the calculations necessary in order for this
scenario to be possible. Before we can test and analyze the attack, we must
consider the possible errors in the used dataset. We also perform an analysis
on the difference between the calculated distances, and the values derived
according to the calculations of Dhondt et al. So can we estimate the
effectiveness of the attack a priori to some extent. Only then do we execute
the attack, evaluate the attacks and form meaningful conclusions.

@@

\subsection{GPS faults}
Some visible data of an activity, which can also be found on
Figure~\ref{fig:activity}, are distance, duration, average speed, etc. The
average speed forms the core to the attack which we will describe in this
article. This will be calculated as the total distance divided by the moving
time of the athlete. Fitness trackers receive raw data from the devices. This
data must therefore be processed before it is usefull fo the users. Especially
GPS-data, which can be exposed to a lot of faults and noise. There are three
types of GPS faults, namely \textit{GPS drift}, \textit{GPS signal loss} and
\textit{GPS bounce}. GPS-drift is a phenomenon where a user's GPS location
deviates from the effective location. This may be caused by densely built
environments, and natural factors such as tall trees. GPS bouncing is an
anomaly caused mainly by tall buildings. In this situation, the GPS signal will
bounce in between buildings on its way to the device from the satellite. The
extra delay that the bounces bring along causes the device to mistakenly think
it has traveled some additional distance. The outcome of the trajectory is then
unpredictable, leading to a `cluster' of GPS points. A last incident that can
occur is GPS signal loss. This occurs when the user's signal is lost, and a new
signal is only received at a later time stamp, causing a jump. A second cause
that can lead to signal loss, which especially applies to fitness trackers, is
the ability to pause an activity. When the activity is resumed again, there
will be a jump in GPS locations, which may lead to miscalculation of distance.

@@

Some techniques can be used to improve the overall accuracy of the GPS data,
and so improve the effectiveness of the attack. The main technique used in this
thesis is \textit{smoothing}, and the hypothesis is that fitnessplatforms use
this technique as well. This is a technique to enhance the accuracy and
reliability of GPS data. There are different implementations of GPS smoothing,
but the one used in this context is moving average filtering. This method
calculates an average position by considering a sliding window of the most
recent GPS measurements. By averaging multiple measurements over a certain time
period, the effects of noise and temporary inaccuracies can be mitigated,
resulting in a smoother and more reliable trajectory. The size of the window
can be chosen. The larger the window, the less accurate the trajectory will be,
but the more noise will be countered. This technique is especially useful to
counter GPS bouncing and GPS drift.

@@

    \item Then we can move on to the identification of Entry Gates (E.G.). Entry gates

@@

          smaller than the maximal possible distance to be covered. Similarly, filtering

@@

\subsection{Geographical distribution}
A geographical distribution is visible in Figure~\ref{fig:heatmap}. It clearly
shows that most activities are located in Central Europe. Additionally, there
is a noticeable concentration in the United States. To a lesser extent, there
are also activities in Australia and South America. The dataset exhibits a
relatively broad spread of activities worldwide, providing a solid foundation
for testing the attack. However, it is important to note that the fraction of
the dataset we have access to, with 101 users, is relatively small, which may
result in a distorted representation of reality.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.85\linewidth]{fig/Afwijkingen&Analyses/Heatmap.png}
    \caption{Geo-heatmap of the users in the dataset}\label{fig:heatmap}
\end{figure}

\subsection{Users and activities}
Table~\ref{tab:stats_dataset} displays several global statistics regarding
users and their associated activities in the dataset. Figure
~\ref{fig:cdf_amount_activities} shows the CDF plot, illustrating the number of
activities per user. It is notable that there is a significant number of
activities available per user in the dataset. However, it's important to note
that the dataset, with an average of 411 activities per user, is not entirely
representative of reality. When comparing these figures with data from a study
conducted by Strava itself in 2020, there seems to be a mismatch.
\begin{table}[h]
    \centering
    \begin{tabular}{|l||c|}
        \hline
                                             & \textbf{Amount} \\
        \hline \hline
        Total \# users                       & 101             \\
        \hline
        Total \# activities                  & 41 554          \\
        \hline
        Average \# activities per user       & 411             \\
        \hline
        Median of the \# activities per user & 296             \\
        \hline
        Maximal \# activities for a user     & 2946            \\
        \hline
        Minimaal \# activities for a user    & 31              \\
        \hline
    \end{tabular}
    \captionsetup{justification=centering}
    \caption{Overview of users and activities}\label{tab:stats_dataset}
\end{table}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\linewidth]{fig/Afwijkingen&Analyses/CDF_amountActivities.jpg}
    \caption{CDF plot of the amount of activities per user}\label{fig:cdf_amount_activities}
\end{figure}

@@

The \textit{Certainty metric} quantifies the concentration of the probability
distribution. A higher Certainty value indicates that the nodes in the
probability distribution are spread further apart, indicating a greater level
of certainty or confidence in the predictions. Essentially, it measures the
degree of dispersion or concentration of the predicted locations in the
probability distribution.

\textit{Spatial Certainty} is a metric that considers the density or concentration of
neighboring nodes in the vicinity of each predicted node, instead of focusing
solely on the probability of each individual node. It takes into account the
density or clustering of predicted locations in the neighbourhood of each node.
A higher Spatial Certainty value indicates a higher density or concentration of
predicted nodes in the local vicinity, suggesting a higher level of certainty
in the predicted locations within their respective neighbourhoods.
